Hands on Lab 2-3: Combining Data from Multiple Sources

### 2018-09-07 DATA607/Hands on Lab 2-3.Rmd
More work on Handon lab 2-3

First I will extract the TB cases from the MySQL database.  I will make use of a MySQL Settings file which hold authentication credentials.  It is not included in the [GitHub repo](https://github.com/mikeasilva/CUNY-SPS/tree/master/DATA607) for obvious reasons.

@@

>>>>

### Step 1-A: Extract the TB Cases From MySQL Database

First I will extract the TB cases from the MySQL database.  I will make use of a MySQL Settings file which hold authentication credentials.  It is not included in the [GitHub repo](https://github.com/mikeasilva/CUNY-SPS/tree/master/DATA607) for obvious reasons.  I will start by connecting to the database.

@@

Now that we have a connection I will pull the total cases by country and year from the database.

```{r pull_data}
sql <- 'SELECT 
    `country`, `year`, SUM(child + adult + elderly) AS `cases`
FROM
    `tb`
GROUP BY `country`, `year`
HAVING `cases` IS NOT NULL'

tb <- dbGetQuery(conn, sql)

# Disconnect from the database
disconnected <- dbDisconnect(conn)

head(tb)
```

### Step 1-B: Retreive the Population From GitHub

I will now bring in the population data from GitHub.

```{r}
url <- 'https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/population.csv'

pop <- read.csv(url)

head(pop)
```

## Step 2: Wrangle the Datasets

The next step is to wrangle the data prior to merging the two datasets together.  I will be joining on the country name and year.  Since there may be variations in the capitalization between the two datasets I will be capitalizing all of them.

The data from the MySQL is not encoded in a way that plays nicely with the population data, so an additional step is taken.

```{r wrangle}
tb <- tb %>%
  mutate(country = enc2utf8(country)) %>%
  mutate(country = toupper(country))

df <- pop %>%
  mutate(country = toupper(country))  %>%
  merge(tb)
```

Since there are `nrow(tb)` observations in the TB dataset and `nrow(df)` in the combined dataset I want to see what fell out.

```{r}
tb_countries <- unique(tb$country)
df_countires <- unique(df$country)
missing <- tb_countries[!tb_countries %in% df_countires]
missing
```

The Ivory Coast fell out of the dataset.
